---
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
---

# Pstat 131 Hw 2

### Noe Arambula ID: 3561131

### 4/5/22

## Libraries

```{r}
# install.packages('ggthemes')
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
tidymodels_prefer()
library(readr)
abalone <- read_csv("abalone.csv")
```

# Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no age variable in the data set. Add age to the data set.

Assess and describe the distribution of age.

```{r}
new_abalone = abalone %>%
  mutate(age = (rings + 1.5))

  ## Distribution of age
ggplot(new_abalone, aes(x = age)) + geom_histogram(color = 'blue', bins = 30)

```

The distribution of age is a little positively skewed most of the points are centered around age 11

# Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

Remember that you'll need to set a seed at the beginning of the document to reproduce your results.

```{r}
set.seed(777)

abalone_split <- initial_split(new_abalone, prop = 0.80,
                                strata = age)  
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```

# Question 3

Using the training data, create a recipe predicting the outcome variable, age, with all other predictor variables. Note that you should not include rings to predict age. Explain why you shouldn't use rings to predict age.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

-   type and shucked_weight,

-   longest_shell and diameter,

-   shucked_weight and shell_weight

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the tidymodels documentation to find the appropriate step functions to use.

```{r}
simple_abalone_recipe = recipe(age ~ type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight + shell_weight, data = abalone_train)

simple_abalone_recipe

  ## we did not include rings to predict age because age is just rings + 1.5 this would scew the data because we would be able to predict age at 100% using rings. We would explain 100% of the variability (R^2) of age by rings

abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight + shell_weight, data = abalone_train) %>% 
  step_dummy(all_nominal_predictors()) %>%   # creates dummy variables
  step_normalize(all_predictors()) %>%    # centers and scales all predictors
  step_interact(terms = type ~ shucked_weight) %>%
  step_interact(terms = longest_shell ~ diameter) %>%
  step_interact(terms = shucked_weight ~ shell_weight)
  
abalone_recipe
```

# Question 4

Create and store a linear regression object using the "lm" engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

# Question 5

Now:

1.  set up an empty workflow,

2.  add the model you created in Question 4, and

3.  add the recipe that you created in Question 3.

```{r}

lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)

lm_fit <- fit(lm_wflow, abalone_train)
head(lm_fit)
```

# Question 6

Use your fit() object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
 # variables represents the new data we will be using to predict an age
variables = data.frame(type = 'F',longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)
predict(lm_fit, new_data = variables)
```

# Question 7

Now you want to assess your model's performance. To do this, use the yardstick package:

1.  Create a metric set that includes R2, RMSE (root mean squared error), and MAE (mean absolute error).

2.  Use predict() and bind_cols() to create a tibble of your model's predicted values from the training data along with the actual observed ages (these are needed to assess your model's performance).

3.  Finally, apply your metric set to the tibble, report the results, and interpret the R2 value.

```{r}
  ## Creating a metric set
abalone_metrics = metric_set(rmse, rsq, mae)

  ## My predictions of the data
abalone_pred <- predict(lm_fit, new_data = abalone_train %>% select(-age))

abalone_pred = bind_cols(abalone_pred, abalone_train %>% select(age))
  
abalone_pred %>%
  head()

  ## apply your metric set to the tibble, report the results, and interpret the R^2 value.
abalone_metrics(abalone_pred, truth = age, estimate = .pred)
  
```

The r-squared value is the goodness of fit measure and measures the relationship between my model and the dependent variable, in this case age.Here the r-squared value is 0.548 for my model meaning 54.8% of the variance of age was explained by my model. This suggests that this model was not that great at estimating values, although still some correlation.
